manuscript_outline

### 2025-05-23 docs/manuscript/manuscript_outline.Rmd
build out example table

There are numerous applications for effective value models and poikilotherms that expand well beyond the bounds of this manuscript. We provide XXX examples in the table below of studies that could be used to parameterize custom models using the `fit_model()` function from hatchR. These are a non-exhaustive search of the both peer-reviewed and grey literature, but demonstrate the wide taxonomic spread that could be paired with the effective value approach.

@@

ggplot() +
  geom_point(data = temperature_data, aes(x = dates, y = temperatures)) +
  geom_line(data = temperature_data, aes(x = dates, y = temperatures)) +
  geom_rect(data = froggy_hatch, aes(xmin = start, xmax = stop, ymax =index-.15, ymin = index-.5, fill = species)) +
  geom_label(data = froggy_hatch, aes(x = start + (stop - start) / 1.45, y = (index -0.325), label = days_to_develop)) +
  labs(x = "Date", y = "Temperature (°C)",
       title = "Development periods of four North American Frog Species") +
  scale_fill_manual(values = c("darkolivegreen4","deepskyblue4", "grey23",  "purple4"), name = "Species") +
  theme_classic() +
  theme(legend.position = c(0.75, 0.25))

@@

>>>>

There are numerous applications for effective value models and poikilotherms that expand well beyond the bounds of this manuscript. We provide XXX examples in the table below of studies that could be used to parameterize custom models using the `fit_model()` function from hatchR. These are a non-exhaustive search of the both peer-reviewed and grey literature, but demonstrate the wide taxonomic breadth that could be paired with the effective value approach.

| Class | Order | Genera | Species | Study |
|:----------------------|:------------|:------------|:------------|:-----------|
| Amphibia | Anura | *Lithobates* | *L. sylvaticus* | @moore1939 |
|  |  |  | *L. pipiens* |  |
|  |  |  | *L. clamitans* |  |
|  |  |  | *L. palustris* |  |
|  |  | *Ascaphus* | *A. truei* | @brown1975 |
|  | Urodela | *Ambystoma* | *A. gracile* | @brown1976 |
| Reptilia | Squamata | *Sceloporus* | *S. undulatus* |  @angilletta2000 |
|  |  | Podarcis | *P. mural*is | @vandamme1992 |
|  | Testudines | Mauremys | M. reevesii | @du2007 |
|  |  |  | 181 species | 141 studies in @while2018 |
| Insecta | Plecoptera | *Nemurella* | *N. pictetii* | @brittain1978, @elliott1984 |
|  |  | *Capnia* | *C. atra* | @brittain1984 |
|  |  | *Capnia* | *C. bifrons* | @elliott1986 |
|  |  | *Mesocapnia* | *M. oenone* | @brittain1984 |
|  |  | *Taeniopteryx* | *T. nebulosa* | @brittain1977 |
| Malacostraca | Decapoda | *Pontastacus* | *P. leptodactylus* | @aydin2004 |

@while2018 not vetted for completeness

@pritchard1987 contains examples of power equations for 18 insect species

```{r, eval=FALSE}

# trying to find study with multiple good pops (failing)

reptiles <- read_csv("../../data/reptile_database.csv")

unique(reptiles$trait)

#species count
reptiles |> 
  filter(trait == "Incubation duration") |> 
group_by(first_author_surname, pub_year, genus, species) |> 
  summarise(count = n())

#study count
reptiles |> 
  filter(trait == "Incubation duration") |> 
group_by(first_author_surname, pub_year) |> 
  summarise(count = n())

#species count
reptiles |> 
  filter(trait == "Incubation duration") |> 
group_by(first_author_surname, pub_year, genus, species, population) |> 
  slice_head() |> 
  group_by(genus, species, state_province) |>  
  summarise(count = n()) |> 
  arrange(desc(count))

# nost common species

reptiles |> 
  filter(trait == "Incubation duration") |> 
  filter( genus == "Caretta" & species == "caretta") |> 
  group_by(first_author_surname, pub_year, population ) |> 
  ggplot(aes(x = T, y = mean, color = first_author_surname, shape = population)) +
  geom_point() +
  geom_line() +
  theme_classic()

reptiles |> 
  filter(trait == "Incubation duration") |> 
  filter( genus == "Sceloporus" & species == "undulatus") |> 
  group_by(first_author_surname, pub_year, population ) |> 
  ggplot(aes(x = T, y = mean, color = first_author_surname, shape = population, group = population)) +
  geom_point() +
  geom_line() +
  theme_classic()
```

examples where other stages could be modeled assuming constant feeding [@lillehammer1986]

@@

p2.1 <- ggplot() +
  geom_point(data = temperature_data, aes(x = dates, y = temperatures)) +
  geom_line(data = temperature_data, aes(x = dates, y = temperatures)) +
  geom_rect(data = froggy_hatch, aes(xmin = start, xmax = stop, ymax =index-.15, ymin = index-.5, fill = species)) +
  geom_label(data = froggy_hatch, aes(x = start + (stop - start) / 1.45, y = (index -0.325), label = days_to_develop)) +
  labs(x = "Date", y = "Temperature (°C)",
       title = "Development periods of four North American Frog Species") +
  scale_fill_manual(values = c("darkolivegreen4","deepskyblue4", "grey23",  "purple4"), name = "Species") +
  theme_classic() +
  theme(legend.position = c(0.75, 0.25))

@@

-   easy to parameterize custom models for quick developing species
